---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# footprint

<!-- badges: start -->
<!-- badges: end -->

The goal of footprint bin point data at very high resolution, storing a sparse grid. These tools use spatstat and raster packages. 

It just exists as the code used in abstraction from the work done in 2015. I'd be happy to collab to extend this to ew work. 

## Installation

You can install the development version of footprint from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("hypertidy/footprint")
```


We create a massive raster, an a fair umber of pretty short line segments, then bin those lines and record cell hits
of the massive raster (we convert from child raster touched by the line segment). The output is 'cell', a pretty massive data frame of `cell,prob`(ability). 

Then, we bin the count of lowest resolution pixels touched by a line into a much lower resolution raster and plot it with the lines. 

It still takes a while, this is like 20min or something so we could parallelize it and perhaps a few other speedups. 



```{r}
library(footprint)  
library(terra)
library(spatstat.geom)
library(dplyr)
llproj <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0"
proj <- "+proj=laea +lat_0=-90 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"

## very round-a-bout, but works
ex <- ext(project(rast(ext(100, 120, -60, -50), nrow = 100, ncol = 100, crs = llproj), proj))

## parent grid (sparse massive grid with 15m pixels)
g <- buildparent(ex, 15)

print(g)  ## don't ever populate the data of this object :)

n <- 1e4
pts0 <- cbind(runif(n, xmin(g) + 3000 , xmax(g) - 3000), runif(n,ymin(g) + 3000, ymax(g) - 3000))
pts1 <- cbind(rnorm(n, pts0[,1], 500), rnorm(n, pts0[,2], 1500))

psegs <- function(x1, x2, add = FALSE) {
  if (!add) plot(rbind(x1, x2), type = "n")
  segments(x1[,1], x1[,2], x2[,1], x2[,2])
}
psegs(pts0, pts1)

diam <- rep(30, nrow(pts0))  ## diameter for density (can be per row )

Kcell <- vector("list", nrow(pts1))
kde <- FALSE

## this could be made a lot faster now 2023-08-07
## --parallelize the loop
## -- (possibly rasterize with terra, can't do kde)
## -- probably can rewrite a fair bit anyway
# for (i in seq(nrow(pts1))) {
#   linp <- bpsp(pts0[i,], pts1[i,], g, diam[i]/2)
#   if (kde) {
#    pix <- density(linp, sigma = diam[i])
#   } else {
#    pix <- pixellate(linp)
#   }
#   rd <- rast(pix)
#   rd[rd < quantile(values(rd)[,1], 0.75)] <- NA_real_
#   
#   vals <- values(rd)
#   Kcell[[i]] <- tibble::tibble(cell = cellFromXY(g, xyFromCell(rd, seq_len(ncell(rd)))[!is.na(vals), ]), 
#         prob = na.omit(vals))  
# if (i %% 50 == 0) print(i)
# }

## this fuction is the old loop above in parallel, which was much harder to do when it was written
## this document is run from the shell via Rscript -e 'rmarkdown::render("README.Rmd")' &
## because, note limitations on future::multicore when run in RStudio
fun_pti <- function(i) {
   linp <- bpsp(pts0[i,], pts1[i,], g, diam[i]/2)
  if (kde) {
   pix <- density(linp, sigma = diam[i])
  } else {
   pix <- pixellate(linp)
  }
  rd <- rast(pix)
  ## values() clashes with future::values 
  rd[rd < quantile(terra::values(rd)[,1], 0.75)] <- NA_real_
  
  vals <- terra::values(rd)
  tibble::tibble(cell = cellFromXY(g, xyFromCell(rd, seq_len(ncell(rd)))[!is.na(vals), ]), 
        prob = na.omit(vals)) 
}

library(furrr)
plan(multicore)
Kcell <- future_map(seq_len(nrow(pts1)), fun_pti)
plan(sequential)

## summarize with dplyr
cell <- do.call(bind_rows, Kcell)

ss <- cell %>% group_by(cell) %>% summarize(prob = sum(prob))

rsum <- setValues(rast(ext(g), res = 50000, crs = crs(g)), 0)
ss$foreign <- cellFromXY(rsum, xyFromCell(g, ss$cell))
xsum <- ss %>% group_by(foreign) %>% summarize(prob = sum(prob)) %>% filter(!is.na(foreign))
rsum[xsum$foreign] <- xsum$prob

plot(rsum)
psegs(pts0, pts1, add = TRUE)
ex2 <- ext(3064556.06339375, 3203002.1878131, -925364.448947992, -816064.877037982)
plot(ex2, add = TRUE)
## zoom in
plot(crop(rsum, ex2))
psegs(pts0, pts1, add = TRUE)

```


## Code of Conduct
  
Please note that the footprint project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

